Seniority and hierarchy configuration interaction for radicals and excited states

### 2023-06-02 Manuscript/manuscript.tex
saving work

Seniority based methods.
Limited studies for radicals and excited states

Hierarchy CI \cite{Kossoski_2022}
Similarly, only for closed-shell systems.

We introduce single-reference CI approximations built on top of an open-shell reference.
Seniority and hCI.
How do these methods perform for ground state radicals?
Impact of perturbative correction.

Ground-state based CI vs. state-specific CI \cite{Kossoski_2023}
Promising results for state-specific excitation-based CI.
We introduce analogous state-specific seniority-based CI ($\Delta$sCI) and hierarchy-based CI ($\Delta$hCI).
$\Delta$sCI2 
$\Delta$hCI1
How do these methods perform for excited states, from both closed-shell and open-shell systems?

@@

%restricted open-shell HF orbitals

>>>>

Configuration interaction (CI) offers a systematic way to solve the many-body electronic structure problem.
By including progressively more determinants in the Hilbert space, the wave function becomes increasingly closer to the exact one, and so does the electronic energy.
In full CI (FCI), all determinants are accounted for and the problem is solved exactly (for a given basis set).
In practice, however, one resorts to approximate CI methods, where only the determinants that satisfy a given criterion are included in the truncated Hilbert space.
For small systems, FCI is still attainable and one can gauge it against approximate CI methods to assesss their performance.

Excitation-based CI
Seniority based CI (sCI)
Limited studies for radicals and excited states
We have recently introduced hierarchy CI (hCI), \cite{Kossoski_2022}
where the Hilbert space is partitioned according to a hierarchy parameter $h$ that combines the excitation degree $e$ and the seniority number $s$.
For different properties and molecular systems, hCI showed an improved convergence to the FCI results when compared to the traditional excitation-based CI. \cite{Kossoski_2022}
However, hCI has only been defined for a closed-shell reference, thus systems with an even number of electrons.
A generalization of the method for radicals has not yet been proposed.

The higher-lying CI roots give access to excited states.
Whereas the performance of excitation-based CI for excited states is well stablished, \cite{} this remains an open question for sCI methods,
despite the growing number of methods targeting excited states that exploit the concept of seniority number. \cite{} 
Similarly, the favourable results of hCI when compared to excitation-based CI for ground states raises the question on how well it can describe excitation energies.

When targeting excited states with CI methods, one can employ the ground-state HF orbitals or instead optimize the orbitals for each state of interest, in a so-called state-specific approach.
Promising results for state-specific excitation-based CI.
\cite{Kossoski_2023}
Here we introduce analogous state-specific seniority-based CI ($\Delta$sCI) and hierarchy-based CI ($\Delta$hCI).
How do these methods perform for excited states, from both closed-shell and open-shell systems?

Here we address the knowledge gaps outlined above by setting four goals.
Our first goal is to generalize hCI \cite{Kossoski_2022} for the case of an open-shell reference determinant, thus extending its domain to radical species.
The second is to assess the performance of this method for ground state radicals, similarly to what has been done for closed-shell systems. \cite{Kossoski_2022}
%Seniority hCI
The third goal is to assess the improvement from perturbative corrections to ground-state radicals and closed-shell systems, for both excitation-based and hierarchy-based CI methods.
Finally, we introduce and gauge seniority-based and hierarchy-based CI methods for excited states, along both the standard ground-state based route and the state-specific route. \cite{Kossoski_2023}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Hierarchy configuration interaction}
\label{sec:hCI}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

state-specific methods
$\Delta$sCI2 
$\Delta$hCI1
%restricted open-shell HF orbitals
